---
title: "Linear Programming"
author: "Matthew Onimus"
date: "4/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lpSolveAPI)
```

## Exercises 1 and 2


```{r exercise1}

exer1 <- make.lp(0,2)

lp.control(exer1, sense = 'max')

set.objfn(exer1, c(1*500, 2*300))
add.constraint(exer1, c(200, 100), "<=", 1200)
add.constraint(exer1, c(1, 2), "<=", 12)
add.constraint(exer1, c(1, 1), "<=", 10)

exer1

solve(exer1)

get.objective(exer1)

get.variables(exer1)

```

```{r exercise2}

exer2 <- make.lp(0,2)

lp.control(exer2, sense = 'max')

set.objfn(exer2, c(2, 3))
add.constraint(exer2, c(20, 10), "<=", 80)
add.constraint(exer2, c(2, -1), ">=", 0)

exer2

solve(exer2)

get.objective(exer2)

get.variables(exer2)


```

# Assignment

I have been known to play some daily fantsy sports and since I like to select my teams with data, I have build the following script to optimize my lineup.  If you are unfamiliar, in general, you are allotted a fixed amount of money to field a roster.  Each player has a different price depending how their stats and a number of other variables.  It is actually quite easy to build an optimized line up based on average points as demonstrated by the code below.

```{r Assignment}

library(stringr)
library(lpSolve)

data6 <- read_csv('DKSalaries.csv') %>% 
  separate(Name, into = c('First Name', 'Last Name'), sep = " ", extra = 'merge') %>% 
  mutate(Name = paste0(`First Name`, " " ,`Last Name`)) %>% 
  select(ID, Name, Position, AvgPointsPerGame, Salary) %>% 
  mutate(AvgPointsPerGame = ifelse(is.na(AvgPointsPerGame), 0, AvgPointsPerGame)) %>% 
  filter(AvgPointsPerGame > 0) %>%
  mutate(Position = ifelse(Position == 'C', '1B', Position)) 

#Set up Indicies####

P_ind <- which(data6[3]=='SP')
F_ind <- which(data6[3]=="1B")
S_ind <- which(data6[3]=='2B')
SS_ind <- which(data6[3]=='SS')
T_ind <- which(data6[3]=='3B')
OF_ind <- which(data6[3]=='OF')

#Create LineUp 1 Build Matrix, add Contraints, solve lp####

A <- matrix(0, nrow = 13, ncol = nrow(data6))

A[1, P_ind] <- 1
A[2, F_ind] <- 1
A[3, F_ind] <- 1
A[4, S_ind] <- 1
A[5, S_ind] <- 1
A[6, SS_ind] <- 1
A[7, SS_ind] <- 1
A[8, T_ind] <- 1
A[9, T_ind] <- 1
A[10, OF_ind] <- 1
A[11, OF_ind] <- 1
A[12, ] <- 1 # assign all players a 1
A[13, ] <- data6$Salary

objective.in <- data6$AvgPointsPerGame #assign the objective
const.mat <- A # assign the constraint matrix
const.dir <- c("=", #Pitchers
               ">=", #min First/Catcher
               "<=", #max first
               ">=", #min 2nd base
               "<=", #max 2nd base
               ">=", #min 3rd base
               "<=", #max 3rd base
               ">=", #min SS
               "<=", #max SS
               ">=", #min OF
               "<=", #max OF
               "=", #Total Players
               "<=") #Salary
const.rhs <- c(1, #Pitchers
               1, #Min first/catcher
               2, #Max first/catcher
               1, #Min 2nd base
               2, #Max 2nd base
               1, #Min 3rd Base
               2, #Max 3rd base
               1, #Min SS
               2, #Max SS
               3, #Min OF
               4, #Max OF
               9, #Total Players
               35000) #Total Salary

sol_1 <- lp(direction = 'max', objective.in, const.mat, const.dir, const.rhs, all.bin = TRUE)

sol_1

inds <- which(sol_1$solution == 1)

results <- data6[inds, ]

results$lineup <- 1

DT::datatable(results)

```

