---
title: "week3AssignmentHyTesting"
author: "Matthew Onimus"
date: "3/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
packages <- c("tidyverse", "DT")
purrr::walk(packages, library, character.only = TRUE)
```

## PDF Exercise

THis first section will recreate the exercises provided in the pdf.

```{r exercise1}

aTraining <- c(12.9,13.5,12.8,15.6,17.2,19.2,12.6,15.3,14.4,11.3)
bTraining <- c(12.7,13.6,12,15.2,16.8,20,12,15.9,16,11.1)

t.test(aTraining, bTraining, paired = TRUE)

qt(0.975, 9)

bTraining2 <- c(12,12.2,11.2,13,15,15.8,12.2,13.4,12.9,11)

t.test(aTraining, bTraining2, paired = TRUE, alt="less")

```

## Read in Excel and Run t-test

We will need to read in the excel sheet, take a look at it, and then perform a t-test on some of the data.

```{r excel}

excel <- readxl::read_excel('paired t-test_3.xlsx')

#it became clear right away this data needed to be reformatted

excelClean <- tibble(
  pt = as.numeric(excel[3, ]),
  stBP = as.numeric(excel[4, ]),
  xyc = as.numeric(excel[5, ]),
  abc = as.numeric(excel[11, ])
) 

excelClean <- excelClean[-1, ]  

xycT <- t.test(excelClean$stBP, excelClean$xyc, paired = TRUE)

xycT

abcT <- t.test(excelClean$stBP, excelClean$abc, paired = TRUE)

abcT


```

Based on the t-test results, the xyc treatment, with a p-value of `r xycT$p.value` is not significant.  These means there is no difference between the means, i.e. the treatment did not help the blood pressure.  The abc treatment option, with a p-value of `r abcT$p.value` is significant.  This indicates that the abc treatment does influence a patients blood pressure.
