---
title: "Week7 Assignment - Unsupervised Learning"
author: "Matthew Onimus"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
library(DT)
library(patchwork)

set.seed(2)
```

## Exercise

In the below code I will reproduce the example from the assignment text.

```{r exercise}

km <- kmeans(iris[,1:4], 3, nstart = 25)

km

table(km$cluster, iris$Species)

```

## Assignment

Here we will apply a clustering algorithm to attempt to group baseball players by position based on some fielding stats

```{r assignment}

fielding2000 <- Lahman::Fielding %>% 
  filter(yearID %in% 2000:2015) %>% 
  select(POS, G, GS, InnOuts, PO, A, E, DP)

kmField <- kmeans(fielding2000[,2:8], 7, nstart = 3)

kmField

table(kmField$cluster, fielding2000$POS)

fieldKM <- fielding2000 %>% 
  bind_cols(cluster = kmField$cluster) %>% 
  mutate(cluster = as.factor(cluster))

a <- ggplot(fieldKM, aes(x = InnOuts, y= PO, color = cluster)) +
  geom_point() +
  theme_classic() +
  labs(
    title = "K-Means Prediction of Position"
  )

b <- ggplot(fieldKM, aes(x = InnOuts, y = PO, color = POS)) +
  geom_point() +
  theme_classic() +
  labs(
    title = "Actual Position of Player"
  )

a/b

```

As you can see, the algorithm had a really though time of doing the prediction based upon PO and InnOuts.  These data are probably going to be really hard to predict by a Kmeans algorithm.  We will try one more model using a PCA

```{r PCA}

fieldingPCA <- prcomp(fielding2000[,2:8])

summary(fieldingPCA)

biplot(fieldingPCA)

```

PCA no better, this data and the vairables I selected just is not good enough to predict position.  Probably not a surprise at all.